Probability of causal contact between interstellar civilizations through MonteCarlo simulations

### 2020-05-13 submit_v03/manuscript.tex
soft biblio

\newcommand{\ceti}{CCN}
\newcommand{\cetis}{CCNs}

@@

We stress the fact that the only observation that can be stated with certainty is that for the number of years projects for the Search of Extraterrestrial Intelligences (SETI) have been working; we have received no signal using the technical possibilities and the conditions established by them \citep{tarter_search_2001}.

@@

The absence of detection has also motivated alternative ideas for new SETI strategies \citep{forgan_exoplanet_2017, balbi_impact_2018, loeb_eavesdropping_2006, maccone_KLT_2010, tarter_advancing_2009, enriquez_breakthrough_2017, loeb_relative_2016, maccone_SETI_2011, lingam_relative_2019, wright_theGsearch_2015, maccone_SETI_2013, maccone_lognormals_2014, harp_application_2018, forgan_possibility_2013, forgan_galactic_2017, funes_searching_2019}.

@@

The Drake equation is a key tool to organise the discussion about the problem of the (unknown) abundance of civilisations in the Galaxy \citep{hinkel_interdisciplinary_2019}.

@@

However, the uncertainties in the factors of this equation make it less prone to a formal study to define searching strategies or to compute the actual number of extraterrestrial intelligences.

@@

The optimistic estimations from the Drake equation also contrast with the SETI initiatives, noticeable overshadowed because despite decades of effort, not a single evidence for intelligent signals has been detected.

@@

\citet{prantzos_joint_2013} proposes a unified framework for a joint analysis of the Drake equation and the Fermi paradox, concluding that for sufficiently long-lived civilisations, colonisation of the Galaxy is the only reasonable option to gain awareness about other life forms.

@@

\citet{haqq-misra_drake_2017} discuss the dependence of the Drake equation parameters on the spectral type of the host stars and the time since the Galaxy formed and examine trajectories for the emergence of communicative civilisations.

@@

Some modifications to the original idea of the equation have tried to imprint a stochastic nature, to propose a probabilistic approach, or to consider the temporal structure missing in the equation.

@@

Temporal aspects of the distribution of communicating civilisations and their contacts have also been explored by several authors \citep{fogg_temporal_1987, forgan_spatiotemporal_2011, balbi_impact_2018, balb_spatiotemporal_2018, horvat_impact_2011}, as well as efforts on considering the stochastic nature of the Drake equation \citep{glade_stochastic_2011}.

@@

The literature about this topic brings a number of works that consider numerical simulations
\citep{forgan_evaluating_2015, vukotic_grandeur_2016, murante_simulating_2015, forgan_numerical_2009, forgan_galactic_2017, ramirez_new_2017}.

@@

Although this approach does not rely on observed quantities that could be modified with time, it also depend on the definition of a number of parameters which are unknown or uncertain.

@@

\citet{grimaldi_signal_2017} consider a statistical model for the probability of the Earth contacting other intelligent civilisations, taking into account the finite lifetime of signal emitters, and based on the fractional volume occupied by all signals reaching our planet.

@@

@@

While the aforementioned assumptions are basic conditions for most of the stochastic processes observed in nature, we need to make stronger hypotheses related to the nature of the causal contact, which is not necessarily tied to any actual message.
%
The contact has, instead, the meaning of a belonging of a node to the light cone of another one.
%
We are considering messages sent through, for example, electromagnetic radiation, gravitational radiation or neutrino communication \citep{hippke_interstellar_2017, wright_how_2018}, but we set apart messages sent with mechanical means or physical objects \citep[e.g., ][]{Armstrong2013, barlow_galactic_2013}, or through some unknown technology that violates the known laws of physics.
% 
For the communication of \cetis{} through messages sent isotropically, we assume that the capacity to emit signals and to receive signals occur at the same time.
%
This means that the ability to find another civilization develops at the same time that the ability and intention to emit a signal in all directions, intended to be detected by an unknown civilization.
%
Although there are several reasons to think that this could not be the case, at large time scales it can be considered that both abilities occur roughly at coincident epochs.
%
Another essential assumption is that all \cetis{} use the same signal power, so that there is a maximum distance out to which it can be detected.
%
It is worth mentioning that we are considering in our experimental setup a system composed by several emitters and receivers across the galaxy, under the hypothesis proposed before.
%
In such a system we compute probabilities of a \ceti{} making contact with another node, and the results must be interpreted as for a randomly chosen \ceti{}, given certain conditions of homogeneity on an ensemble of equivalent galaxies.
%
Then, we stress the fact that we are not centered on the probabilities for the Earth itself, but on a generalized \ceti{}.
%
Regarding the extent of the signals, we know that the distance from which a signal from Earth could be detected using the current technology is of a few hundreds light years, given that the signal was sent to a specific direction.
%
It is straightforward to propose and implement a distribution of maximum distances, although this would increase the model complexity at the cost of a larger uncertainty or of another dimension on the hypothesis space.

@@

The Galactic Habitable Zone is set at fixed values of the inner and outer
radii, radial symmetry is assumed, and the with of the galactic disk
is neglected compared to the radial size.
%
The two radii are measured in light years, with the aim to maintain a
single and comprehensive unit for both space and time coordinates. 
%
The simulation was implemented on a python 3 code, which is publicly
available\footnote{https://github.com/mlares/simu\_contact}

@@

%\bibliographystyle{mn2e}
\setlength{\bibsep}{0.0pt}

\begin{thebibliography}{90}
\expandafter\ifx\csname natexlab\endcsname\relax\def\natexlab#1{#1}\fi

\bibitem[{Adamic(2000)}]{adamic_zipf_2000}
Adamic L., 2000, {Zipf, Power-laws, and Pareto - a ranking tutorial}. web

\bibitem[{Anchordoqui \& Weber(2019)}]{anchordoqui_upper_2019}
Anchordoqui L.~A., Weber S.~M., 2019, arXiv e-prints

\bibitem[{Annis(1999)}]{annis_astrophysical_1999}
Annis J., 1999, JBIS - Journal of the British Interplanetary Society, 52, 19

\bibitem[{Armstrong \& Sandberg(2013)}]{Armstrong2013}
Armstrong S., Sandberg A., 2013, Acta Astronautica, 89, 1

\bibitem[{Balb(2018)}]{balb_spatiotemporal_2018}
Balb A., 2018, Mem. della Soc. Astron. Ital., 89, 425

\bibitem[{Balbi(2018)}]{balbi_impact_2018}
Balbi A., 2018, Astrobiology, 18, 54

\bibitem[{Barab{\'{a}}si(2009)}]{barabasi_scale_2009}
Barab{\'{a}}si A.~L., 2009, Science (80-. )., 325, 412

\bibitem[{Barlow(2013{\natexlab{a}})}]{barlow_galactic_2012}
Barlow M.~T., 2013{\natexlab{a}}, Int. J. Astrobiol., 12, 63

\bibitem[{Barlow(2013{\natexlab{b}})}]{barlow_galactic_2013}
Barlow M.~T., 2013{\natexlab{b}}, Int. J. Astrobiol., 12, 63

\bibitem[{Benguigui \& Marinov(2016)}]{benguigui_classificacion_2016}
Benguigui L., Marinov M., 2016, {A classification of the natural and social
  distributions Part 2: the explanations}. Tech. rep.

\bibitem[{Bloetscher(2019)}]{bloetscher_using_2019}
Bloetscher F., 2019, Acta Astronaut., 155, 118

\bibitem[{Borra(2012)}]{borra_searching_2012}
Borra E.~F., 2012, Astronomical Journal, 144

\bibitem[{Brin(1983)}]{brin_great_1983}
Brin G., 1983, Quarterly Journal of the Royal Astronomical Society, 24, 283

\bibitem[{Burchell(2006)}]{burchell_whither_2006}
Burchell M.~J., 2006, Int. J. Astrobiol., 5, 243

\bibitem[{Carroll-Nellenback {et~al}\mbox{.}(2019)Carroll-Nellenback, Frank,
  Wright, \& Scharf}]{carroll_nellemback_fermi_2019}
Carroll-Nellenback J., Frank A., Wright J., Scharf C., 2019, The Astronomical
  Journal, 158, 117

\bibitem[{Chung(2003)}]{chung_simulation_2003}
Chung C.~A., 2003, {Simulation Modeling Handbook: A Practical Approach
  (Industrial and Manufacturing Engineering Series)}. CRC press

\bibitem[{{\'{C}}irkovi{\'{c}}(2004)}]{cirkovic_temporal_2004}
{\'{C}}irkovi{\'{c}} M.~M., 2004, Astrobiology, 4, 225

\bibitem[{{Conway Morris}(2018)}]{conway_three_2018}
{Conway Morris} S., 2018, Int. J. Astrobiol., 17, 287

\bibitem[{Dayal {et~al}\mbox{.}(2016)Dayal, Ward, \&
  Cockell}]{dayal_habitability_2016}
Dayal P., Ward M., Cockell C., 2016, arXiv e-prints

\bibitem[{Do{\v{s}}ovi{\'{c}} {et~al}\mbox{.}(2019)Do{\v{s}}ovi{\'{c}},
  Vukoti{\'{c}}, \& {\'{C}}irkovi{\'{c}}}]{Dosovic2019}
Do{\v{s}}ovi{\'{c}} V., Vukoti{\'{c}} B., {\'{C}}irkovi{\'{c}} M.~M., 2019,
  Astronomy and Astrophysics, 625

\bibitem[{Drake(1962)}]{drake_intelligent_1962}
Drake F., 1962

\bibitem[{Edmondson \& Stevens(2003)}]{Edmondson2003}
Edmondson W.~H., Stevens I.~R., 2003, International Journal of Astrobiology, 2,
  231

\bibitem[{Enriquez {et~al}\mbox{.}(2017)Enriquez, Siemion, Foster, Gajjar,
  Hellbourg, Hickish, Isaacson, Price, Croft, DeBoer, Lebofsky, MacMahon, \&
  Werthimer}]{enriquez_breakthrough_2017}
Enriquez J.~E. {et~al.}, 2017, Astrophys. J., 849, 104

\bibitem[{Fogg(1987)}]{fogg_temporal_1987}
Fogg M.~J., 1987, Icarus, 69, 370

\bibitem[{Forgan {et~al}\mbox{.}(2016)Forgan, Dayal, Cockell, \&
  Libeskind}]{forgan_evaluating_2015}
Forgan D., Dayal P., Cockell C., Libeskind N., 2016, Int. J. Astrobiol., 16, 60

\bibitem[{Forgan(2009)}]{forgan_numerical_2009}
Forgan D.~H., 2009, Int. J. Astrobiol., 8, 121

\bibitem[{Forgan(2011)}]{forgan_spatiotemporal_2011}
Forgan D.~H., 2011, Int. J. Astrobiol., 10, 341

\bibitem[{Forgan(2013)}]{forgan_possibility_2013}
Forgan D.~H., 2013, JBIS - J. Br. Interplanet. Soc., 66, 144

\bibitem[{Forgan(2014)}]{forgan_collimated_2014}
Forgan D.~H., 2014, JBIS - J. Br. Interplanet. Soc., 67, 232

\bibitem[{Forgan(2017{\natexlab{a}})}]{forgan_galactic_2016}
Forgan D.~H., 2017{\natexlab{a}}, Int. J. Astrobiol., 16, 349

\bibitem[{Forgan(2017{\natexlab{b}})}]{forgan_galactic_2017}
Forgan D.~H., 2017{\natexlab{b}}, Int. J. Astrobiol., 16, 349

\bibitem[{Forgan(2019)}]{forgan_exoplanet_2017}
Forgan D.~H., 2019, Int. J. Astrobiol., 18, 189

\bibitem[{Forgan \& Rice(2010)}]{forgan_numerical_2010}
Forgan D.~H., Rice K., 2010, Int. J. Astrobiol., 9, 73

\bibitem[{Frank(2009)}]{frank_common_2009}
Frank S.~A., 2009, {The common patterns of nature}

\bibitem[{Funes {et~al}\mbox{.}(2019)Funes, Florio, Lares, \&
  Asla}]{funes_searching_2019}
Funes J.~G., Florio L., Lares M., Asla M., 2019, Theol. Sci., 0, 1

\bibitem[{Galera {et~al}\mbox{.}(2019)Galera, Galanti, \&
  Kinouchi}]{galera_invasion_2019}
Galera E., Galanti G.~R., Kinouchi O., 2019, International Journal of
  Astrobiology, 18, 316

\bibitem[{Glade {et~al}\mbox{.}(2012)Glade, Ballet, \&
  Bastien}]{glade_stochastic_2011}
Glade N., Ballet P., Bastien O., 2012, Int. J. Astrobiol., 11, 103

\bibitem[{Gobat \& Hong(2016)}]{gobat_evolution_2016}
Gobat R., Hong S.~E., 2016, Astron. Astrophys., 592

\bibitem[{Gonzalez(2005)}]{gonzalez_habitable_2005}
Gonzalez G., 2005, Orig. Life Evol. Biosph., 35, 555

\bibitem[{Gonzalez {et~al}\mbox{.}(2001)Gonzalez, Brownlee, \&
  Ward}]{gonzalez_galactic_2001}
Gonzalez G., Brownlee D., Ward P., 2001, Icarus, 152, 185

\bibitem[{Grimaldi(2017)}]{grimaldi_signal_2017}
Grimaldi C., 2017, Sci. Rep., 7, 46273

\bibitem[{Grimaldi {et~al}\mbox{.}(2018)Grimaldi, Marcy, Tellis, \&
  Drake}]{Grimaldi2018}
Grimaldi C., Marcy G.~W., Tellis N.~K., Drake F., 2018, Publications of the
  Astronomical Society of the Pacific, 130

\bibitem[{Haqq-Misra(2019)}]{haqq-misra_evolution_2019}
Haqq-Misra J., 2019, arXiv e-prints

\bibitem[{Haqq-Misra \& Kopparapu(2018)}]{haqq-misra_drake_2017}
Haqq-Misra J., Kopparapu R.~K., 2018, Habitability of the Universe Before
  Earth, 307

\bibitem[{Harp {et~al}\mbox{.}(2018)Harp, Ackermann, Astorga, Arbunich,
  Barrios, Hightower, Meitzner, Barott, Nolan, Messerschmitt, Vakoch, Shostak,
  \& Tarter}]{harp_application_2018}
Harp G.~R. {et~al.}, 2018, Astrophys. J., 869, 66

\bibitem[{Hart(1975)}]{hart_explanation_1975}
Hart M.~H., 1975, Quarterly Journal of the Royal Astronomical Society, 16, 128

\bibitem[{Hinkel {et~al}\mbox{.}(2019)Hinkel, Hartnett, Lisse, \&
  Young}]{hinkel_interdisciplinary_2019}
Hinkel N., Hartnett H., Lisse C., Young P., 2019, Bull. Am. Astron. Soc., 51,
  497

\bibitem[{Hippke(2017)}]{hippke_interstellar_2017}
Hippke M., 2017, arXiv e-prints

\bibitem[{Horvat(2006)}]{horvat_calculating_2007}
Horvat M., 2006, Int. J. Astrobiol., 5, 143

\bibitem[{Horvat {et~al}\mbox{.}(2011)Horvat, Naki{\'{c}}, \&
  Oto{\v{c}}an}]{horvat_impact_2011}
Horvat M., Naki{\'{c}} A., Oto{\v{c}}an I., 2011, Int. J. Astrobiol., 11, 51

\bibitem[{Jeong {et~al}\mbox{.}(2000)Jeong, Tombor, Albert, Oltvai, \&
  Barabasi}]{jeong_large_2000}
Jeong H., Tombor B., Albert R., Oltvai Z.~N., Barabasi A.~L., 2000, Nature,
  407, 651

\bibitem[{Lampton(2013)}]{lampton_information_2013}
Lampton M., 2013, Int. J. Astrobiol., 12, 312

\bibitem[{Lineweaver {et~al}\mbox{.}(2004)Lineweaver, Fenner, \&
  Gibson}]{lineweaver_galactic_2004}
Lineweaver C.~H., Fenner Y., Gibson B.~K., 2004, Science (80-. )., 303, 59

\bibitem[{Lingam \& Loeb(2018)}]{lingam_relative_2019}
Lingam M., Loeb A., 2018, Astrobiology, 19, 28

\bibitem[{Loeb {et~al}\mbox{.}(2016)Loeb, Batista, \&
  Sloan}]{loeb_relative_2016}
Loeb A., Batista R.~A., Sloan D., 2016, J. Cosmol. Astropart. Phys., 2016, 040

\bibitem[{Loeb \& Zaldarriaga(2007)}]{loeb_eavesdropping_2006}
Loeb A., Zaldarriaga M., 2007, J. Cosmol. Astropart. Phys., 1

\bibitem[{Maccone(2010)}]{maccone_KLT_2010}
Maccone C., 2010, Acta Astronaut., 67, 1427

\bibitem[{Maccone(2011)}]{maccone_mathematical_2011}
Maccone C., 2011, Orig. Life Evol. Biosph., 41, 609

\bibitem[{MacCone(2011)}]{maccone_SETI_2011}
MacCone C., 2011, Acta Astronaut., 68, 63

\bibitem[{MacCone(2013)}]{maccone_SETI_2013}
MacCone C., 2013, Int. J. Astrobiol., 12, 218

\bibitem[{Maccone(2014{\natexlab{a}})}]{maccone_evolution_2014}
Maccone C., 2014{\natexlab{a}}, Int. J. Astrobiol., 13, 290

\bibitem[{Maccone(2014{\natexlab{b}})}]{maccone_lognormals_2014}
Maccone C., 2014{\natexlab{b}}, Acta Astronaut., 105, 538

\bibitem[{Maccone(2015)}]{maccone_statistical_2015}
Maccone C., 2015, Acta Astronaut., 115, 277

\bibitem[{Mart{\'{i}}n \& Goldenfeld(2006)}]{martin_origin_2006}
Mart{\'{i}}n H.~G., Goldenfeld N., 2006, Proc. Natl. Acad. Sci. U. S. A., 103,
  10310

\bibitem[{Mitzenmacher(2004)}]{mitzenmacher_brief_2004}
Mitzenmacher M., 2004, Internet Math., 1, 226

\bibitem[{Morrison \& Gowanlock(2015)}]{morrison_extending_2015}
Morrison I.~S., Gowanlock M.~G., 2015, Astrobiology, 15, 683

\bibitem[{Murante {et~al}\mbox{.}(2015)Murante, Monaco, Borgani, Tornatore,
  Dolag, \& Goz}]{murante_simulating_2015}
Murante G., Monaco P., Borgani S., Tornatore L., Dolag K., Goz D., 2015, Mon.
  Not. R. Astron. Soc., 447, 178

\bibitem[{Newman(2005)}]{newman_power_2005}
Newman M., 2005, Contemp. Phys., 46, 323

\bibitem[{Newman \& Sagan(1981)}]{newman_galactic_1981}
Newman W.~I., Sagan C., 1981, Icarus, 46, 293

\bibitem[{Peters(2018)}]{peters_outer_2018}
Peters T., 2018, Int. J. Astrobiol., 17, 282

\bibitem[{Prantzos(2013)}]{prantzos_joint_2013}
Prantzos N., 2013, Int. J. Astrobiol., 12, 246

\bibitem[{Ptolemaeus(2014)}]{ptolemaeus_system_2014}
Ptolemaeus C., 2014, {System Design, Modeling, and Simulation. Using Ptolemy
  II}. Ptolemy.org, Berkeley, p. 690

\bibitem[{Rahvar(2017)}]{rahvar_cosmic_2016}
Rahvar S., 2017, Mon. Not. R. Astron. Soc., 470, 3095

\bibitem[{Ramirez {et~al}\mbox{.}(2017)Ramirez, G{\'{o}}mez-Mu{\~{n}}oz,
  V{\'{a}}zquez, \& N{\'{u}}{\~{n}}ez}]{ramirez_new_2017}
Ramirez R., G{\'{o}}mez-Mu{\~{n}}oz M.~A., V{\'{a}}zquez R., N{\'{u}}{\~{n}}ez
  P.~G., 2017, International Journal of Astrobiology, 1

\bibitem[{Ross(2012)}]{ross_simulation_2012}
Ross S.~M., 2012, {Simulation}. Elsevier Science Publishing Co Inc

\bibitem[{Simkin \& Roychowdhury(2006)}]{simkin_theory_2006}
Simkin M.~V., Roychowdhury V.~P., 2006, J. Math. Sociol., 30, 33

\bibitem[{Smith(2009)}]{smith_broadcasting_2009}
Smith R.~D., 2009, Int. J. Astrobiol., 8, 101

\bibitem[{Solomonides {et~al}\mbox{.}(2016)Solomonides, Kaltenegger, \&
  Terzian}]{solomonides_probabilistic_2016}
Solomonides E., Kaltenegger L., Terzian Y., 2016, 228th AAS, San Diego, 228, 1

\bibitem[{Sornette(2006)}]{sornette_critical_2006}
Sornette D., 2006, {Critical phenomena in natural sciences: Chaos, fractals,
  selforganization and disorder: Concepts and tools}. Springer

\bibitem[{Sotos(2019)}]{Sotos_biotechnology_2019}
Sotos J.~G., 2019, {Biotechnology and the lifetime of technical civilizations}

\bibitem[{Starling \& Forgan(2014)}]{starling_virulence_2013}
Starling J., Forgan D.~H., 2014, Int. J. Astrobiol., 13, 45

\bibitem[{Tarter(2001)}]{tarter_search_2001}
Tarter J., 2001, Annu. Rev. Astron. Astrophys., 39, 511

\bibitem[{Tarter {et~al}\mbox{.}(2009)Tarter, Backus, Blair, Cordes, Harp,
  Henry, Horowitz, Howard, Kilsdonk, Korpela, Lasio, Levin, Shostack, \&
  Werthimer}]{tarter_advancing_2009}
Tarter J. {et~al.}, 2009, Astro2010 Astron. Astrophys. Decad. Surv. Sci. White
  Pap. no. 294, 2010

\bibitem[{Vakoch \& Dowd(2015)}]{vakoch_drake_2015}
Vakoch D.~A., Dowd M.~F., 2015, {The drake equation: Estimating the prevalence
  of extraterrestrial life through the ages}, Vakoch D.~A., Dowd M.~F., eds.
  Cambridge University Press, Cambridge, pp. 1--319

\bibitem[{VanHouten(2018)}]{vanhouten_isthere_2017}
VanHouten M.~A., 2018, Lancet Gastroenterol. Hepatol., 3, 382

\bibitem[{Vukoti{\'{c}} \&
  {\'{C}}irkovi{\'{c}}(2012)}]{vukotic_astrobiological_2012}
Vukoti{\'{c}} B., {\'{C}}irkovi{\'{c}} M.~M., 2012, Orig. Life Evol. Biosph.,
  42, 347

\bibitem[{Vukoti{\'{c}} {et~al}\mbox{.}(2016)Vukoti{\'{c}}, Steinhauser,
  Martinez-Aviles, {\'{C}}irkovi{\'{c}}, Micic, \&
  Schindler}]{vukotic_grandeur_2016}
Vukoti{\'{c}} B., Steinhauser D., Martinez-Aviles G., {\'{C}}irkovi{\'{c}}
  M.~M., Micic M., Schindler S., 2016, Mon. Not. R. Astron. Soc., 459, 3512

\bibitem[{Walters {et~al}\mbox{.}(1980)Walters, Hoover, \&
  Kotra}]{walters_interstellar_1980}
Walters C., Hoover R.~A., Kotra R., 1980, Icarus, 41, 193

\bibitem[{Wright {et~al}\mbox{.}(2015)Wright, Cartier, Zhao, Jontof-Hutter, \&
  Ford}]{wright_theGsearch_2015}
Wright J.~T., Cartier K. M.~S., Zhao M., Jontof-Hutter D., Ford E.~B., 2015,
  Astrophys. J., 816, 17

\bibitem[{Wright {et~al}\mbox{.}(2018)Wright, Kanodia, \&
  Lubar}]{wright_how_2018}
Wright J.~T., Kanodia S., Lubar E., 2018, The Astronomical Journal, 156, 260

\end{thebibliography}

>>>>

\newcommand{\ceti}{CCN}
\newcommand{\cetis}{CCNs}
\newcommand{\hs}{\textsc{hearsay}}

@@

We stress the fact that the only observation that can be stated with
certainty is that for the number of years projects for the Search of
Extraterrestrial Intelligences (SETI) have been working; we have
received no signal using the technical possibilities and the
conditions established by them \citep{tarter_search_2001}.

@@

The absence of detection has also motivated alternative ideas for new
SETI strategies \citep{forgan_exoplanet_2017, balbi_impact_2018,
loeb_eavesdropping_2006, maccone_KLT_2010, tarter_advancing_2009,
enriquez_breakthrough_2017, loeb_relative_2016, maccone_SETI_2011,
lingam_relative_2019, wright_theGsearch_2015, maccone_SETI_2013,
maccone_lognormals_2014, harp_application_2018,
forgan_possibility_2013, forgan_galactic_2017, funes_searching_2019}.

@@

The Drake equation is a key tool to organise the discussion about the
problem of the (unknown) abundance of civilisations in the Galaxy
\citep{hinkel_interdisciplinary_2019}.

@@

However, the uncertainties in the factors of this equation make it
less prone to a formal study to define searching strategies or to
compute the actual number of extraterrestrial intelligences.

@@

The optimistic estimations from the Drake equation also contrast with
the SETI initiatives, noticeable overshadowed because despite decades
of effort, not a single evidence for intelligent signals has been
detected.

@@

\citet{prantzos_joint_2013} proposes a unified framework for a joint
analysis of the Drake equation and the Fermi paradox, concluding that
for sufficiently long-lived civilisations, colonisation of the Galaxy
is the only reasonable option to gain awareness about other life
forms.

@@

\citet{haqq-misra_drake_2017} discuss the dependence of the Drake
equation parameters on the spectral type of the host stars and the
time since the Galaxy formed and examine trajectories for the
emergence of communicative civilisations.

@@

Some modifications to the original idea of the equation have tried to
imprint a stochastic nature, to propose a probabilistic approach, or
to consider the temporal structure missing in the equation.

@@

Temporal aspects of the distribution of communicating civilisations
and their contacts have also been explored by several authors
\citep{fogg_temporal_1987, forgan_spatiotemporal_2011,
balbi_impact_2018, balb_spatiotemporal_2018, horvat_impact_2011}, as
well as efforts on considering the stochastic nature of the Drake
equation \citep{glade_stochastic_2011}.

@@

The literature about this topic brings a number of works that consider
numerical simulations \citep{forgan_evaluating_2015,
vukotic_grandeur_2016, murante_simulating_2015, forgan_numerical_2009,
forgan_galactic_2017, ramirez_new_2017}.

@@

Although this approach does not rely on observed quantities that could
be modified with time, it also depend on the definition of a number of
parameters which are unknown or uncertain.

@@

\citet{grimaldi_signal_2017} consider a statistical model for the
probability of the Earth contacting other intelligent civilisations,
taking into account the finite lifetime of signal emitters, and based
on the fractional volume occupied by all signals reaching our planet.

@@

We need to raise the scope of our results by stressing the fact that we are considering causal contacts instead of actual contacts.
%
By a causal contact we mean a configuration, in space and time, in which a node belongs to the light cone of another node.
%
This means that a causal contact does not need to be tied to any actual message emission or detection.
%
In more realistic scenarios, there are several sources of ‘’signal loss’’ with respect to the ideal case.
%
Among them, we can mention temporal and signal power aspects and direction dependent communication capabilties.
%
The results must then be interpreted as the case of ideal nodes, with a perfect efficiency in the emision and reception of signals.
%
The probabilities of contacts presented here are then upper limits to the number of communications between civilization s in the Galaxy, given the implemented hipotheses in the model.
%
The use of light cones as causal contact regions is inspired by the fact that light—speed traveling messengers like 
electromagnetic radiation, gravitational radiation or neutrinos are often considered as possible message carriers \citep{hippke_interstellar_2017, wright_how_2018}.
%
This excludes messages sent with mechanical means or physical objects \citep[e.g., ][]{Armstrong2013, barlow_galactic_2013}, or through some unknown technology that violates the known laws of physics.

As part of this benchmark, we assume that the capacity to emit and receive signals occur at the same time.
%
Although there are several reasons to think that this could not be the case, at large time scales it can be considered that both abilities occur roughly at coincident epochs.
%
In the ideal setup, this would be equivalent to nodes that send messages isotropically and scan the local skies on all directions with a perfect eficiency.
%
Another essential assumption is that all \cetis{} use the same signal power, so that there is a maximum distance out to which it can be detected.
%
% It is worth mentioning that we are considering in our experimental setup a system composed by several emitters and 
% receivers across the galaxy, under the hypothesis proposed before.
%
In such a system we compute probabilities of a \ceti{} making contact with another node, and the results must be interpreted as for a randomly chosen \ceti{}, given certain conditions of homogeneity on an ensemble of equivalent galaxies.
%
Then, we stress the fact that we are not centred on the probabilities for the Earth itself, but on a generalised \ceti{}. 
%
Regarding the extent of the signals, we know that the distance from which a signal from Earth could be detected using the current technology is of a few hundreds light years, given that the signal was sent to a specific direction.
%
It is straightforward to propose and implement a distribution of maximum distances, although this would increase the model complexity at the cost of a larger uncertainty or of another dimension on the hypothesis space.

% For the communication of \cetis{} through messages sent isotropically, we assume that the capacity to emit signals and %to receive signals occur at the same time. This means that the ability to find another civilization develops at the same %time that the ability and intention to emit a signal in all directions, intended to be detected by an unknown civilization.

@@

@@

The Galactic Habitable Zone is set at fixed values of the inner and outer
radii, radial symmetry is assumed, and the with of the galactic disk
is neglected compared to the radial size.
%
The two radii are measured in light years, with the aim to maintain a
single and comprehensive unit for both space and time coordinates. 
%
%}}}

\subsection{Model implementation}
%{{{

% hearsay 
%
With the aim of making a reproducible project, we developed the tools
that allow to run and obtain the results shown in this work.
%
The simulations were implemented on a Python-3 code, dubbed \hs{},
which is publicly accessible through the GitHub
platform\footnote{https://github.com/mlares/simu\_contact} under the
MIT-license.
%
The project is in the process of registration with the ‘’Astrophysics
Source Code Library’’ \citep[ASCL, ][]{2015JORS....3E..15A,
2020ASPC..522..731A}
%
In order to achieve high performance we made use of vectorised methods
and array broadcasting properties provided by the NumPy package
\citep{oliphant_numpy_2006, vanderwalt_numpy_2011}
%
The combination of C-compiled Numpy routines for numerical
computations with a high-level language allows to obtain an easily
comprehensible and maintainable code.
%
From the viewpoint of a user, \hs{} is an object-oriented package that
exposes the main functionalities as classes and methods.
%
In the configuration step, the user prepares the set of simulation
parameters through an initialisation file.
%
Since configuration files and simulation results are persistent, it is
straightforward to keep track of different experiments and the
experiments can be revisited easily.
%
An in-depth description of the methods can be found in the
documentation, which is automatically generated from \hs{} docstrings
and made public in the read-the-docs service.

%Quality assurance
%
Regarding the quality assurance of the software, we endorse the PEP-8
style guide for python (van Rossum et al. 2001) throughout the entire
project, which improves the code maintainability and helps to minimise
the so called ''code-entropy'' of future versions.
%
The deviations in style were analysed with the flake8 tool, which
improves the readability and consistency of the code.
%
Testing and code coverage have been applied to the package and the
development process as standard qualitative and quantitative metrics.
%
The different parts and functions can be validated to assert if they
return the expected results on specific test cases and testing
scenarios \citep{jazayeri_engeneering_2007}.
%
On the other hand, code coverage measures the fraction visited and
executed in the code by the testing procedure
\citep{miller_testing_1963}.
%
The advantage of an exhaustive code-coverage is that it ensures that
most of the sources of errors have been considered.
%
In the \hs{} project we provide test tools for the distinct parts of
the simulation setup and run, reaching a 99\% of code coverage in the
version used for this work.
%
For every new version of the code, all these quality assurance methods
and metrics are run automatically by a continuous-integration service.
%
It is possible to reproduce the experiments performed in this work
using the \hs{} package.
%
The installation is available through a git repository and also on the
Python Package Index (PyPI).
%
Since the simulation setup is configurable, the time required for a
simulation to complete depends on the simulation parameters.

@@

\bibliographystyle{mn2e}
\setlength{\bibsep}{0.0pt}
\bibliography{biblio_seti, biblio_software}
